---
title: "Day 10 Afternoon Practice"
format: html
---

```{r}
library(tidyverse)
library(here)
library(kableExtra)
```


```{r}
# load in data 
ice_mads <- read_csv(here("data", "ntl33_v7.csv"), na = c("-999", "NA"))
meterological <- read_csv(here("data", "ntl20_v6.csv"))

# data looks good. in tidy form already 
```


- Create visualization that compares ice duration across different lakes in Madison Lake Area. 
- Include both the actual ice duration values and some representation of summary statistics for each lake. 
From metadata 
- ME = Lake Mendota 
- MO = Lake Monona 
- WI = Lake Wingra


```{r}
# data viz comparing ice_duration and lakeid
# remove nas from ice_duration 

ice_duration_rmna <- filter(ice_mads, !is.na(ice_duration)) %>%
  mutate(lake_name = )
  
  
ggplot(data = ice_mads, aes(x = lakeid, y = ice_duration)) +
  geom_boxplot() +
  labs(y = "Ice duration (days)",
       x = "Lake ID",
       title = "Lakes of Madison Area Ice Duration")


ice_duration_rmna %>% 
  group_by(lakeid) %>% 
  summarise(mean = mean(ice_duration),
            sd = sd(ice_duration),
            sample = n()) %>% 
  kable(col.names = c("Lake ID", "Mean", "Standard Deviation", "Sample Size")) %>% 
  kable_classic() %>% 
  add_header_above(c("ME = Lake Mendota\nMO = Lake Monona\nWI = Lake Wingra\n" = 1, " " = 1, " " = 2)) 
```

- Find the mean ice cover duration in Madison Lake Area by year
```{r}
ice_duration_rmna %>% 
  group_by(year4) %>%
  summarise(mean_ice = mean(ice_duration)) %>% 
  kable(col.names = c("Year", "Mean ice duration (days)")) %>% 
  kable_paper() %>% 
  scroll_box(height = "150px", width = "500px")
```


- Create an exploratory visualization of mean ice cover duration by year for the Madison Lake Area.
```{r}
mean_ice_cover_year <- ice_duration_rmna %>% 
  group_by(year4) %>%
  summarise(mean_ice = mean(ice_duration))

ggplot(data = mean_ice_cover_year, aes(x = year4, y = mean_ice)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Year",
       y = "Mean ice cover duration (days)")
```
The overall trend of mean ice cover duration is that there is a stead decrease in mean duration from the years 1850 to the year 2019. I included a trend line in my plot to show this pattern. 



- Find the mean air temperature (using the daily average air temperature - adjusted) in Madison Lake Area by year using only observations from winter months (December, January, February)
```{r}
#mean ave_air_temp_adjusted grouped by year only including the months dec(12), jan(1), feb(2) 

mean_air_temp <- meterological %>% 
  filter(month == 12 | month == 1 | month == 2) %>% 
  group_by(year4, month) %>% 
  summarise(mean_air = mean(ave_air_temp_adjusted))
```


